---
title: "Oil & Gas Wells' Petrophysical and Geographical Data"
author: "Quynh Nguyen (qmn76), Mark Metzger (mm73828), Hanna Jones ()"
date: "October 26, 2016"
output: 
  html_document:
    toc: true
---

### Session Info

```{r}
sessionInfo()
```

### Processing Data

#### STEP 1: Find a dataset. 

The data set used is extracted from a CSV file called "GASISData.csv" data contains data from GASIS Reservoir Data System. GASIS, or Gas Information System, is an initiative by The Department of Energy to "combine previously compiled public domain data, newly acquired and interpreted data, and newly released data into a single database with over 18,000 reservoir records."

The data set contains comprehensive field/reservoir identification, location, reservoir type, parameters, etc for oil and gas wells across the U.S. and is a useful tool to analyze the wells' petrophysical and geographical data on a large scale.

```{r}
read.csv("../01 Data/GASISData.csv", stringsAsFactors = FALSE)
```

#### STEP 2: Clean up the data Using Extract, Transform, and Load techniques.

Below is a subset of our data. The data set we used is a table of different oil & gas reservoirs across different states, geological regions, and geological era. In addition, it also contains different sets of the porosity and permeability values. 

```{r}
source("../01 Data/Data.R", echo = TRUE)
```

### Getting Started With Shiny App 

#### STEP 3: Creating A Histogram Plotter

#### STEP 4: Creating A K-means Clustering Plotter

#### STEP 5: Creating Something (by Hanna)

#### STEP 6: Creating A Widgets To Filter Data

