---
title: "Oil & Gas Wells' Petrophysical and Geographical Data"
author: "Quynh Nguyen (qmn76), Mark Metzger (mm73828), Hanna Jones (hmj427)"
date: "October 26, 2016"
output: 
  html_document:
    toc: true
---

## Session Info

```{r}
sessionInfo()
```

## Finding & Processing Data

### STEP 1: Find a dataset. 

The data set used is extracted from a CSV file called "GASISData.csv" data contains data from GASIS Reservoir Data System. GASIS, or Gas Information System, is an initiative by The Department of Energy to "combine previously compiled public domain data, newly acquired and interpreted data, and newly released data into a single database with over 18,000 reservoir records."

The data set contains comprehensive field/reservoir identification, location, reservoir type, parameters, etc for oil and gas wells across the U.S. and is a useful tool to analyze the wells' petrophysical and geographical data on a large scale.

Column Name | Column Name Explanation
------------- | -------------
STATE | state name
FLDNAME | DOE/EIA field name 
RESNAME | reservoir name
R_STUDY | reservoir source (Atlas, Reserv. Study, TOTL)
PLAYNAME | gas atlas geologic play name 
PLAYCOD | gas atlas geologic play code
ERANM | geologic era name
SUBPLAYN | gas atlas subplay name
SYSNM | geologic system name
SERNM | geologic series name
S_POR | data source for porosity 
S_PERM | data source for permeability 
LINKA | unique key for reservoir data system 
AVPOR | average porosity of pay interval 
LOPOR | porosity range minimum -net pay
HIPOR | porosity range maximum -net pay 
AVPERM | average permeability of pay interval 
LOPERM | permeability range minimum 
HIPERM | permeability range maximum 
RESTEMP | average reservoir temperature 

```{r}
read.csv("../01 Data/GASISData.csv", stringsAsFactors = FALSE)
```

### STEP 2: Clean up the data Using Extract, Transform, and Load techniques.

Below is a subset of our data. The data set we used is a table of different oil & gas reservoirs across different states, geological regions, and geological era. In addition, it also contains different sets of the porosity and permeability values. 


### STEP 3: Grabbing DataFrame from OracleDB.

```{r}
require("jsonlite")
require("RCurl")

df <- data.frame(fromJSON(getURL(URLencode('oraclerest.cs.utexas.edu:5001/rest/native/?query="select * from GASISDATA"'),httpheader=c(DB='jdbc:oracle:thin:@aevum.cs.utexas.edu:1521/f16pdb', USER='cs329e_qmn76', PASS='orcl_qmn76', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE)))
```


## Viewing Data

### Summary of Data

```{r}
summary(df)
```

### Preview of Data
```{r}
head(df)
```

### Subset of Data

Below is a subset of data where the HIPERM (or High Permeability) is greater than .5.

```{r}
subset <- subset(df, HIPERM > .5)
head(subset)
```

Here's another example of the data where SYSNM (or geologic system name) is CRETACEOUS.

```{r}
subset <- subset(df, SYSNM == 'CRETACEOUS')
head(subset)
```


## Visualizing Data with Shiny App 

First we need to load the neccessary packages.

On the UI Side, we need shiny and shinydashboard.

```{r}
library(shiny)
require(shinydashboard)
```

On the server side, we need other packages that would allow us to read in data from the Oracle SQL database, process it, and plot diferent visualizations.

```{r}
require("jsonlite")
require("RCurl")
require(ggplot2)
require(dplyr)
require(shiny)
```

### STEP 4: Creating A Histogram Plotter

In order to create a histogram plotter, we will create a list of options on the sidebar panel and plot the result in the main panel.

In this case, we will let the users decide which column of data they want to plot and how many cells they want to split into. The final result will find the frequency, or count of times a number within a range appears.

Interesting, we can see that some petrophysical variables like Reservoir Temperature follows a normal distribution while other follows almost a bi-normal distribituion.

### STEP 5: Creating A K-means Clustering Plotter

In order to create a K-means clustering plotter, we will create a list of x and y options, the number of clusters, etc. on the sidebar panel and plot the K-means clustering result in the main panel. 

The final diagram plots the x and y values as scatter points and then try to do k-means analysis. K-means clustering is a method that attemps to divide the dataset into k number of clusters by connecting each data point to its closest mean value.

### STEP 6: Creating a Box Plot

We thought it would be interesting to visualize the difference in reservoir temperatures for various dimensions. Specifically, this visualization aims to answer the following: How do reservoir temperatures differ between states, geological eras, and geologic systems?

To do this, we first filtered our data by adjusting the SQL query to exclude any null reservoir temperatures.

We created an option to exclude or include outliers, as they may skew the graph and thus our conclusions about what the graph says. We also created a list of dimensions for which we would like to use to compare temperatures, and displayed this in a drop down for the user to select.

While this graph does not show a detailed explanation of temperatures, it allows for a high-level visualization of the disparity in different states and geological eras or systems.

### STEP 7: Creating A Widgets To Filter Data

We organized our app into a ShinyUI application with a sidebar panel where we placed filters for the user. An example of a filter we created is below:


This allows the user to select an input from the list for the Box Plot that we created. When the user selects a choice, the graph updates accordingly.

In the main panel, we placed all of the graphs that we created in the steps above. An example of how this is done is below:

### Putting it All Together

Below is the full server code and UI code. The server code includes all of our database queries and creation of the plots. The UI code is how we organized the Shiny application and created widgets for user input.

**[Final Result](https://cs329equynhmarkhanna.shinyapps.io/02Shiny/)**


